He-Man: A domain-specific language for lightweight concurrency

### 2012-03-06 paper/paper.tex
some stuff; going to change decls

\documentclass[preprint,10pt]{sigplanconf}
\usepackage[colorlinks=true,linkcolor=black,urlcolor=black,citecolor=black]%
{hyperref}
\usepackage{graphicx}
\usepackage{verbatim}

@@

\begin{figure}[ht]
\centering
\begin{code}
child_code = declare_thread [("child_fd", Int)] $
  \[child_fd] -> do
  ev <- setup_connection child_fd
  buf <- new_buf bufsize
  while 1 $ do
    amt_read <- do_read ev buf bufsize
    ifthen (amt_read .== 0) exit
    amt_written <- full_write ev buf amt_read
    ifthen (amt_written .< amt_read) exit
\end{code}
\caption{He-Man example from echo server.}
\label{fig:code-ex}
\end{figure}

@@

We offload the work of type checking He-Man programs to the Haskell
type checker. Our method for doing this is fairly standard, following
Leijen and Meijer \cite{Leijen:1999:DSE:1267936.1267945} in using
phantom types to enforce typing.

@@

TODO: Talk about argument packets.

\begin{figure}[ht]
\centering
\begin{code}
class ArgPacket a b | a -> b, b -> a where
  toIExprList :: a -> [IExpr]
  makeVars :: [String] -> a
class ArgDecls a b | a -> b, b -> a where
  toVDecl :: a -> [VDecl]

@@

instance ArgDecls (TVDecl a) [a] where
  toVDecl (x1, t1) = [(x1, toIType t1)]
instance ArgDecls (TVDecl a, TVDecl b) (a, b) where
  toVDecl ((x1, t1), (x2, t2)) =
    [(x1, toIType t1), (x2, toIType t2)]
-- ...

@@

The per-thread memory overhead imposed by He-Man is very small: it
consists of a thread id, list pointers for the scheduling queue, the
next state to enter, a pointer to the most recently received event,
and list heads and tails for lists of the events being waited on,
events owned by the thread, and buffers owned by the thread (the
latter two are so that we can automatically reclaim these resources
when the thread exits). On a 64-bit system, the total overhead comes
out to 88 bytes, and could easily be reduced further (e.g., by using
singly-linked lists).  For comparison, the default thread stack size
under NPTL on Linux is 2 megabytes.

>>>>

\documentclass[preprint]{sigplanconf}
\usepackage[colorlinks=true,linkcolor=black,urlcolor=black,citecolor=black]%
{hyperref}
\usepackage{graphicx}
\usepackage{verbatim}

@@

\begin{figure}[ht]
\centering
\begin{code}
child_code = declare_thread (Int) $
  \[child_fd] -> do
  ev <- setup_connection child_fd
  buf <- new_buf bufsize
  while 1 $ do
    amt_read <- do_read ev buf bufsize
    ifthen (amt_read .== 0) exit
    amt_written <- full_write ev buf amt_read
    ifthen (amt_written .< amt_read) exit
\end{code}
\caption{He-Man example from echo server.}
\label{fig:code-ex}
\end{figure}

@@

We offload the work of type checking He-Man programs to the Haskell type
checker. Our method for doing this is fairly standard, following Leijen and
Meijer \cite{LeijenMeijer} in using phantom types to enforce typing.

@@

Type checking foreign function calls and thread declaration and
spawning is somewhat more interesting.
% and maybe a bit novel? I dunno.
For foreign function calls, our main concern was being able to write a
single wrapper for call that handles a varying number of typed
arguments. To this end, we created an \t{ArgPacket} type class that
requires the ability to convert a packet into a list of \t{IExpr} and
to construct an \t{ArgPacket} from a list of variable names. Instances
of \t{ArgPacket} are provided for packets of $0$ to $5$ \t{Expr}
arguments. Now we can parameterize a \t{Call} wrapper over an
\t{ArgPacket} to provide a \t{Expr} based interface.

For thread declarations, we can produce a similar \t{ArgDecl} class
and associated instances that can convert a packet of \t{Type}s into a
list of \t{IType}s.

We want more than this, however: we want to ensure that both the code
for a declared thread uses its arguments properly at the declared
types code and that spawning threads provides arguments of the correct
types. To do this, we need some way to extract all of the phantom
types of the \t{Expr}s and \t{Type}s in \t{ArgPacket}s and
\t{ArgDecl}s

\begin{figure}[ht]
\centering
\begin{code}
class ArgPacket a b | a -> b, b -> a where
  toIExprList :: a -> [IExpr]
  makeVars :: [String] -> a
class ArgDecls a b | a -> b, b -> a where
  toITypeList :: a -> [IType]

@@

instance ArgDecls (Type a) [a] where
  toITypeList t1 = [toIType t1]
instance ArgDecls (Type a, Type b) (a, b) where
  toITypeList (t1, t2) = [toIType t1, toIType t2]
-- ...

@@

The per-thread memory overhead imposed by He-Man is very small: it
consists of a thread id, list pointers for the scheduling queue, the
next state to enter, a pointer to the most recently received event,
and list heads and tails for lists of the events being waited on,
events owned by the thread, and buffers owned by the thread (the
latter two are so that we can automatically reclaim these resources
when the thread exits). On a 64-bit system, the total overhead comes
out to 88 bytes, and could easily be reduced further (e.g., by using
singly-linked lists).  For comparison, the default thread stack size
under NPTL on Linux is 2 megabytes.
%% XXX update with new info?
